Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-25 paper.tex
typo: subject verb agreement
adding some description

\newcommand{\TOOdyn}[0]{\star}
\newcommand{\TOOpre}[1]{#1}
\newcommand{\POOunit}[0]{\mathtt{Unit}}
\newcommand{\POOfun}[2]{#1 \shortrightarrow #2}
\newcommand{\POOprod}[2]{#1 \times #2}
\newcommand{\POOsum}[2]{#1 \plus #2}
\newcommand{\eOOvar}[1]{#1}
\newcommand{\eOOsole}[0]{\mathtt{unit}}
\newcommand{\eOOlam}[4]{\lambda^{#1\rightarrow{}#2}#3.#4}
\newcommand{\eOOapp}[2]{(#1 \; #2)}
\newcommand{\eOOcons}[4]{\mathtt{cons}^{\POOprod{#1}{#2}} \; #3 \; #4}
\newcommand{\eOOcar}[1]{\mathtt{fst} \; #1}
\newcommand{\eOOcdr}[1]{\mathtt{snd} \; #1}
\newcommand{\eOOinl}[3]{\mathtt{inl}^{\POOsum{#1}{#2}} \; #3}
\newcommand{\eOOinr}[3]{\mathtt{inr}^{\POOsum{#1}{#2}} \; #3}
\newcommand{\eOOcase}[3]{\mathtt{case} \; #1 \; #2 \; #3}
\newcommand{\eOOcast}[4]{#1 \langle \cOOcast{#2}{#3}{#4} \rangle}
\newcommand{\eOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\cOOcast}[3]{#1 \Rightarrow^{#2} #3}
\newcommand{\oOOinj}{\mathtt{dyn}}
\newcommand{\oOOsole}{\mathtt{unit}}
\newcommand{\oOOfun}{\mathtt{fun}}
\newcommand{\oOOcons}{\mathtt{cons}}
\newcommand{\oOOinl}{\mathtt{inl}}
\newcommand{\oOOinr}{\mathtt{inr}}
\newcommand{\oOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\vOOcast}[2]{#1\langle#2\rangle}
\newcommand{\vOOfun}[3]{\mathtt{fun} \; #1 \; #2 \; #3}
\newcommand{\vOOtt}[0]{\mathtt{unit}}
\newcommand{\vOOcons}[2]{\mathtt{cons}\;#1\;#2}
\newcommand{\vOOinl}[1]{\mathtt{inl}\;#1}
\newcommand{\vOOinr}[1]{\mathtt{inr}\;#1}
\newcommand{\rOOsucc}[1]{\mathtt{succ}\;#1}
\newcommand{\rOOfail}[1]{\mathtt{fail}\;#1}

@@

\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #2 \; #1 \; #3 \; #4 \; #5}

@@

The syntax and static semantics is the same for the \lazyD{} and
\lazyUD{} cast calculi and is reviewed in \figref{fig:blame-static}.
As usual, the important features are the cast expressions,
$\eOOcast{e}{T_1}{l}{T_2}$, which are responsible for runtime type
checking, and blame expressions, $\eOOblame{l}$, that raise errors.
The syntax and static semantics is the same that of
\citet{siek2009exploring} except for a few minor exceptions:
\begin{itemize}
\item We add sum, product, and unit types.
\item We separate types into those with a type constructor at the top,
  the \emph{pretypes} ($\POOunit$, functions, products, and sums),
  versus the dynamic type $\star$ (a.k.a. $\mathtt{Dyn}$ or
  $\mathbb{?}$).
\item We annotate the type of many expressions explicitly because we 
refer to them in the dynamic semantics. These expressions include cons 
expressions, left injections, right injections, and lambda abstractions.
\end{itemize}
As usual, the source $T_1$ and target types $T_2$ of a cast
$\eOOcast{e}{T_1}{l}{T_2}$ are required to be consistent, written $T_1
\sim T_2$. The consistency relation is standard and defined in
\figref{fig:blame-static}.

@@

  Reduction \fbox{$\judgeCreduce{s}{s}$}
  \[
  \begin{array}{rclr}
    \hiredrule{
    \sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{
      \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa}
%     \langle\cOOcast{T_1}{l}{T_2}\rangle\kappa
    }
  }{}
  \redrule{
    \sOOinspect{\eOOvar{x}}{\mathcal{E}}{\kappa}
  }{  
    \sOOreturn{\mathcal{E}(x)}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOsole}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\vOOtt}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\vOOfun{\mathcal{E}}{x}{e}}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOcons{T_1}{T_2}{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{\kOOconsI{T_1}{T_2}{e_2}{\mathcal{E}}{\kappa}}
  }{}
  \redrule{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{\kOOinl{T_1}{T_2}{\kappa}}
  }{}
  \redrule{
  \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
  \sOOinspect{e}{\mathcal{E}}{\kOOinr{T_1}{T_2}{\kappa}}
  }{}
  \redrule{
    \sOOinspect{\eOOapp{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
\sOOinspect{e_1}{\mathcal{E}}{\kOOappI{e_2}{\mathcal{E}}{\kappa}}}{}

@@

  Reduction \fbox{$\judgeSreduce{C}{s}{s}$}
  \[
  \begin{array}{rclr}
  \hiredruleS{
    \sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{ext(cast(T_1,l,T_2),\kappa)}
  }{}
  \redruleS{
    \sOOinspect{\eOOvar{x}}{\mathcal{E}}{\kappa}
  }{  
    \sOOreturn{\mathcal{E}(x)}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOsole}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\hcvOOtt}{\kappa}
  }{}
  \hiredruleS{
    \sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\hcvOOfun{id(T_1)}{\mathcal{E}}{x}{e}{id(T_2)}}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOcons{T_1}{T_2}{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{
      wrap(\kOOconsI{T_1}{T_2}{e_2}{\mathcal{E}}{\kappa})
    }
  }{}
  \redruleS{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOinl{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOinr{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOapp{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{wrap(\kOOappI{e_2}{\mathcal{E}}{\kappa})}}{}

@@

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation
$s_1 \approx s_2$, where $s_1 \in \effCEK{C}$ and $s_2 \in \ineffCEKD{}$.
It is the smallest congruence relation that also includes the
pairs induced by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa : T$, where $T$ is their source type, and values, written $v 
\approx v : T$, where $T$ is their type.
%
If $k \approx \kappa : T$, adding an identity cast on top of $k$ gives a 
related continuation. This handles the way in which
\effCEK{C} uses $wrap$ to add identity coercions around continuations
whereas \ineffCEKD{} does not.
%
%\todo[inline]{The use of ``wrap'' obscures this explanation.
%  Consider getting rid of wrap by inlining it. -JS
%\\---\\
%I agree that inlining wrap and ext can improve readability of $\kappa \approx 
%\kappa$, but this would make latex different from Agda, is it alright?
%-KC
%\\---\\
%Yes, that is alright. Inlining is really a trivial change.
%-JS
%}
%
If $\kOOcast{c}{k_1} \approx \kappa_2 : T$, extending them with 
the same cast gives related continuations.
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a naked function in \ineffCEKD, the function in \effCEK{C} 
must use identities as its casts.
%
If functions are related, we can apply the same cast on both sides.
%
Rules for products and sums are similar.

\begin{figure}
%  $\text{\effCEK{C}} \approx \text{\ineffCEKD}$
% \[\dots\]
  \fbox{$s \approx s$}
  % Because the grammar for states of the two machines are not identical,
  % I think it is worthwhile to write out these rules explicitly. -Jeremy
  \begin{gather*}
    \inference{dom(\mathcal{E}_1) = dom(\mathcal{E}_2) \\
               \forall x {:} T \in dom(\mathcal{E}_1),\,
                   \mathcal{E}_1(x) \approx \mathcal{E}_2(x) : T\\
               \kappa_1 \approx \kappa_2 : T}
              {\langle e, \mathcal{E}_1, \kappa_1 \rangle \approx
              \langle e, \mathcal{E}_2, \kappa_2 \rangle}
    %% \quad
    %% \inference{v_1 \approx v_2 : T \quad \kappa_1 \approx \kappa_2 : T}
    %%           {\langle v_1, \kappa_1 \rangle  \approx
    %%           \langle v_2, \kappa_2 \rangle}
    \quad
    \inference{}
              {\sOOhalt{o} \approx \sOOhalt{o}}
  \end{gather*}
  \fbox{$\kappa \approx \kappa : T$}
  \begin{gather*}
  \inference{
    k \approx \kappa : T
  }{
    \kOOcast{id(T)}{k} \approx \kappa
  }
  \quad
  \inference{
    \kOOcast{c}{k_1} \approx \kappa_2 : T_2
  }{
    \kOOcast{seq(cast(T_1,l,T_2),c)}{k_1} \approx 
    \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
  }
\\
  \end{gather*}
\\  
  \fbox{$v \approx v : T$}
  \begin{gather*}
  \inference{
    v_1 \approx v_2 : I
  }{
    \hcvOOinj{I}{v_1} \approx \vOOcast{v_2}{\cOOcast{I}{l}{\TOOdyn}}
    : \TOOdyn
  }
\\ \\
   \inference{}{
    \hcvOOtt \approx \vOOtt : \POOunit
   }
   \quad
  \inference{}{
    \hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
    : \POOunit
  }
  \\ \\
  \inference{
    \mathcal{E}_1 \approx \mathcal{E}_2 &
    dom(\mathcal{E}_1) = dom(\mathcal{E}_2) = \Gamma \\
    \Gamma , x : T_1 \vdash e : T_2
  }{
    \hcvOOfun{id(T_1)}{\mathcal{E}_1}{x}{e}{id(T_2)}
    \approx
    \vOOfun{\mathcal{E}_2}{x}{e}
    : \POOfun{T_1}{T_2}
  }
\\ \\
\inference{
\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2} \\
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
}{
\hcvOOfun{c_1'}{\mathcal{E}}{x}{e}{c_2'}
\approx
\vOOcast{v_2}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\\ \\
  \inference{
    v_1 \approx v_2 : T_1 &
    v_3 \approx v_4 : T_2 &
  }{
    \hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
    \approx
    \vOOcons{v_2}{v_4}
    : \POOprod{T_1}{T_2}
  }
\\ \\
  \inference{
    \hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOprod{T_1}{T_2} \\
    c_1' = seq(c_1,cast(T_1,l,T_3)) \\
    c_2' = seq(c_2,cast(T_2,l,T_4))
  }{
    \hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
    \vOOcast{v_2}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
    : \POOprod{T_3}{T_4}
  }
\\ \\
  \inference{
    v_1 \approx v_2 : T_1
  }{
    \hcvOOinl{v_1}{id(T_1)} \approx \vOOinl{v_2}
    : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
    \hcvOOinl{v_1}{c} \approx v_2
    : \POOsum{T_1}{T_2} &
    c' = seq(c,cast(T_1,l,T_3))
  }{
    \hcvOOinl{v_1}{c'} \approx
    \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
    : \POOsum{T_3}{T_4} 
  }
  \\
  \inference{
  v_1 \approx v_2 : T_2
  }{
  \hcvOOinr{v_1}{id(T_2)} \approx \vOOinr{v_2}
  : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
  \hcvOOinr{v_1}{c} \approx v_2
  : \POOsum{T_1}{T_2} &
  c' = seq(c,cast(T_2,l,T_4))
  }{
  \hcvOOinr{v_1}{c'} \approx
  \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
  : \POOsum{T_3}{T_4}
  }
  \end{gather*}
  \caption{Bisimulation between \effCEK{C} and \ineffCEKD}
  \label{fig:bisim-SC-D}
\end{figure}

>>>>

\newcommand{\TOOdyn}[0]{\star}
\newcommand{\TOOpre}[1]{#1}
\newcommand{\POOunit}[0]{\mathtt{Unit}}
\newcommand{\POOfun}[2]{#1 \shortrightarrow #2}
\newcommand{\POOprod}[2]{#1 \times #2}
\newcommand{\POOsum}[2]{#1 \plus #2}
\newcommand{\eOOvar}[1]{#1}
\newcommand{\eOOsole}[0]{\mathtt{unit}}
\newcommand{\eOOlam}[4]{\lambda^{#1\rightarrow{}#2}#3.#4}
\newcommand{\eOOapp}[2]{(#1 \; #2)}
\newcommand{\eOOcons}[4]{\mathtt{cons}^{\POOprod{#1}{#2}} \; #3 \; #4}
\newcommand{\eOOcar}[1]{\mathtt{fst} \; #1}
\newcommand{\eOOcdr}[1]{\mathtt{snd} \; #1}
\newcommand{\eOOinl}[3]{\mathtt{inl}^{\POOsum{#1}{#2}} \; #3}
\newcommand{\eOOinr}[3]{\mathtt{inr}^{\POOsum{#1}{#2}} \; #3}
\newcommand{\eOOcase}[3]{\mathtt{case} \; #1 \; #2 \; #3}
\newcommand{\eOOcast}[4]{#1 \langle \cOOcast{#2}{#3}{#4} \rangle}
\newcommand{\eOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\cOOcast}[3]{#1 \Rightarrow^{#2} #3}
\newcommand{\oOOinj}{\mathtt{dyn}}
\newcommand{\oOOsole}{\mathtt{unit}}
\newcommand{\oOOfun}{\mathtt{fun}}
\newcommand{\oOOcons}{\mathtt{cons}}
\newcommand{\oOOinl}{\mathtt{inl}}
\newcommand{\oOOinr}{\mathtt{inr}}
\newcommand{\oOOblame}[1]{\mathtt{blame} \; #1}
\newcommand{\vOOcast}[2]{#1\langle#2\rangle}
%\newcommand{\vOOfun}[3]{\mathtt{fun} \; #1 \; #2 \; #3}
% Jeremy prefers the following
\newcommand{\vOOfun}[3]{\langle \lambda  #2. #3, #1 \rangle}
\newcommand{\vOOtt}[0]{\mathtt{unit}}
\newcommand{\vOOcons}[2]{\mathtt{cons}\;#1\;#2}
\newcommand{\vOOinl}[1]{\mathtt{inl}\;#1}
\newcommand{\vOOinr}[1]{\mathtt{inr}\;#1}
\newcommand{\rOOsucc}[1]{\mathtt{succ}\;#1}
\newcommand{\rOOfail}[1]{\mathtt{fail}\;#1}

@@

%\newcommand{\hcvOOfun}[5]{\mathtt{fun} \; #2 \; #1 \; #3 \; #4 \; #5}
% Jeremy prefers the following version:
\newcommand{\hcvOOfun}[5]{\langle \lambda #3.\, #4 , #1 , #5, #2 \rangle}

@@

The syntax and static semantics are the same for the \lazyD{} and
\lazyUD{} cast calculi. They are reviewed in \figref{fig:blame-static}.
As usual, the important features are the cast expressions,
$\eOOcast{e}{T_1}{l}{T_2}$, which are responsible for runtime type
checking, and blame expressions, $\eOOblame{l}$, that raise errors.
The syntax and static semantics is the same that of
\citet{siek2009exploring} except for a few minor exceptions:
\begin{itemize}
\item We add sum, product, and unit types.
\item We separate types into those with a type constructor at the top,
  the \emph{pretypes} ($\POOunit$, functions, products, and sums),
  versus the dynamic type $\star$ (a.k.a. $\mathtt{Dyn}$ or
  $\mathbb{?}$).
\item We annotate the type of many expressions explicitly because we 
refer to them in the dynamic semantics. These expressions include cons 
expressions, left injections, right injections, and lambda abstractions.
\end{itemize}
As usual, the source $T_1$ and target types $T_2$ of a cast
$\eOOcast{e}{T_1}{l}{T_2}$ are required to be consistent, written $T_1
\sim T_2$. The consistency relation is standard and defined in
\figref{fig:blame-static}.

@@

  Transition \fbox{$\judgeCreduce{s}{s}$}
  \[
  \begin{array}{rclr}
    \hiredrule{
    \sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{
      \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa}
%     \langle\cOOcast{T_1}{l}{T_2}\rangle\kappa
    }
  }{}
  \redrule{
    \sOOinspect{\eOOvar{x}}{\mathcal{E}}{\kappa}
  }{  
    \sOOreturn{\mathcal{E}(x)}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOsole}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\vOOtt}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\vOOfun{\mathcal{E}}{x}{e}}{\kappa}
  }{}
  \redrule{
    \sOOinspect{\eOOcons{T_1}{T_2}{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{\kOOconsI{T_1}{T_2}{e_2}{\mathcal{E}}{\kappa}}
  }{}
  \redrule{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{\kOOinl{T_1}{T_2}{\kappa}}
  }{}
  \redrule{
  \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
  \sOOinspect{e}{\mathcal{E}}{\kOOinr{T_1}{T_2}{\kappa}}
  }{}
  \redrule{
    \sOOinspect{\eOOapp{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
\sOOinspect{e_1}{\mathcal{E}}{\kOOappI{e_2}{\mathcal{E}}{\kappa}}}{}

@@

  Transition \fbox{$\judgeSreduce{C}{s}{s}$}
  \[
  \begin{array}{rclr}
  \hiredruleS{
    \sOOinspect{\eOOcast{e}{T_1}{l}{T_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{ext(cast(T_1,l,T_2),\kappa)}
  }{}
  \redruleS{
    \sOOinspect{\eOOvar{x}}{\mathcal{E}}{\kappa}
  }{  
    \sOOreturn{\mathcal{E}(x)}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOsole}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\hcvOOtt}{\kappa}
  }{}
  \hiredruleS{
    \sOOinspect{\eOOlam{T_1}{T_2}{x}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOreturn{\hcvOOfun{id(T_1)}{\mathcal{E}}{x}{e}{id(T_2)}}{\kappa}
  }{}
  \redruleS{
    \sOOinspect{\eOOcons{T_1}{T_2}{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{
      wrap(\kOOconsI{T_1}{T_2}{e_2}{\mathcal{E}}{\kappa})
    }
  }{}
  \redruleS{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOinl{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOinl{T_1}{T_2}{e}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e}{\mathcal{E}}{wrap(\kOOinr{T_1}{T_2}{\kappa})}
  }{}
  \redruleS{
    \sOOinspect{\eOOapp{e_1}{e_2}}{\mathcal{E}}{\kappa}
  }{
    \sOOinspect{e_1}{\mathcal{E}}{wrap(\kOOappI{e_2}{\mathcal{E}}{\kappa})}}{}

@@

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation
$s_1 \approx s_2$, where $s_1 \in \effCEK{C}$ and $s_2 \in \ineffCEKD{}$.
It is the smallest congruence relation that also includes the
pairs induced by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa : T$, where $T$ is their source type, and values, written $v 
\approx v : T$, where $T$ is their type.
%
If $k \approx \kappa : T$, adding an identity cast on top of $k$ gives a 
related continuation. This handles the way in which
\effCEK{C} uses $wrap$ to push identity coercions onto continuations
whereas \ineffCEKD{} does not.
%
We remark that this bisimulation relation is unusual in that it uses
some ADT operations in its definition, such as $id$, $cast$, and
$seq$, which is necessary because \effCEK{C} and the bisimulation
is parameterized over the Cast ADT.
%
%\todo[inline]{The use of ``wrap'' obscures this explanation.
%  Consider getting rid of wrap by inlining it. -JS
%\\---\\
%I agree that inlining wrap and ext can improve readability of $\kappa \approx 
%\kappa$, but this would make latex different from Agda, is it alright?
%-KC
%\\---\\
%Yes, that is alright. Inlining is really a trivial change.
%-JS
%}
%
If $\kOOcast{c}{k_1} \approx \kappa_2 : T$, extending the right-hand
side with a cast $T_1 \Rightarrow^l T_2$ and sequencing
$cast(T_1,l,T_2)$ with $c$ on the left-hand side, produces related
continuations. This takes care of the difference between
the transition rules for $e \langle T_1 \Rightarrow^l T_2$
on the two machines.
%
If two values of type $I$ are related, injecting them to $\TOOdyn$
gives related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a closure in \ineffCEKD, the closure in \effCEK{C} must
have identities as its casts.
%
If a closure on the left is related to a value, we can apply
equivalent casts on both sides to obtain related values.
%
The rules for products and sums are similar.

\begin{figure}
%  $\text{\effCEK{C}} \approx \text{\ineffCEKD}$
% \[\dots\]
  \fbox{$s \approx s$}
  % Because the grammar for states of the two machines are not identical,
  % I think it is worthwhile to write out these rules explicitly. -Jeremy
  \begin{gather*}
    \inference{dom(\mathcal{E}_1) = dom(\mathcal{E}_2) \\
               \forall x {:} T \in dom(\mathcal{E}_1),\,
                   \mathcal{E}_1(x) \approx \mathcal{E}_2(x) : T\\
               \kappa_1 \approx \kappa_2 : T}
              {\langle e, \mathcal{E}_1, \kappa_1 \rangle \approx
              \langle e, \mathcal{E}_2, \kappa_2 \rangle}
    %% \quad
    %% \inference{v_1 \approx v_2 : T \quad \kappa_1 \approx \kappa_2 : T}
    %%           {\langle v_1, \kappa_1 \rangle  \approx
    %%           \langle v_2, \kappa_2 \rangle}
    \quad
    \inference{}
              {\sOOhalt{o} \approx \sOOhalt{o}}
  \end{gather*}
  \fbox{$\kappa \approx \kappa : T$}
  \begin{gather*}
  \inference{
    k \approx \kappa : T
  }{
    \kOOcast{id(T)}{k} \approx \kappa
  }
  \quad
  \inference{
    \kOOcast{c}{k_1} \approx \kappa_2 : T_2
  }{
    \kOOcast{seq(cast(T_1,l,T_2),c)}{k_1} \approx 
    \kOOcast{\cOOcast{T_1}{l}{T_2}}{\kappa_2}
  }
\\
  \end{gather*}
\\  
  \fbox{$v \approx v : T$}
  \begin{gather*}
  \inference{
    v_1 \approx v_2 : I
  }{
    \hcvOOinj{I}{v_1} \approx \vOOcast{v_2}{\cOOcast{I}{l}{\TOOdyn}}
    : \TOOdyn
  }
  \\ \\
   %% This one is an obvious congruence rule. -Jeremy
   %% \inference{}{
   %%  \hcvOOtt \approx \vOOtt : \POOunit
   %% }
   %% \quad
  \inference{}{
    \hcvOOtt \approx \vOOcast{\vOOtt}{\cOOcast{\POOunit}{l}{\POOunit}}
    : \POOunit
  }
  \\ \\
  \inference{
    \mathcal{E}_1 \approx \mathcal{E}_2 &
    dom(\mathcal{E}_1) = dom(\mathcal{E}_2) = \Gamma \\
    \Gamma , x : T_1 \vdash e : T_2
  }{
    \hcvOOfun{id(T_1)}{\mathcal{E}_1}{x}{e}{id(T_2)}
    \approx
    \vOOfun{\mathcal{E}_2}{x}{e}
    : \POOfun{T_1}{T_2}
  }
\\ \\
\inference{
\hcvOOfun{c_1}{\mathcal{E}}{x}{e}{c_2} \approx v_2 : \POOfun{T_1}{T_2} \\
c_1' = seq(cast(T_3,l,T_1),c_1) \\
c_2' = seq(c_2,cast(T_2,l,T_4))
}{
\hcvOOfun{c_1'}{\mathcal{E}}{x}{e}{c_2'}
\approx
\vOOcast{v_2}{\cOOcast{\POOfun{T_1}{T_2}}{l}{\POOfun{T_3}{T_4}}}
: \POOfun{T_3}{T_4}
}
\\ \\
  \inference{
    v_1 \approx v_2 : T_1 &
    v_3 \approx v_4 : T_2 &
  }{
    \hcvOOcons{v_1}{id(T_1)}{v_3}{id(T_2)}
    \approx
    \vOOcons{v_2}{v_4}
    : \POOprod{T_1}{T_2}
  }
\\ \\
  \inference{
    \hcvOOcons{v_1}{c_1}{v_3}{c_2} \approx v_2 : \POOprod{T_1}{T_2} \\
    c_1' = seq(c_1,cast(T_1,l,T_3)) \\
    c_2' = seq(c_2,cast(T_2,l,T_4))
  }{
    \hcvOOcons{v_1}{c_1'}{v_3}{c_2'} \approx
    \vOOcast{v_2}{\cOOcast{\POOprod{T_1}{T_2}}{l}{\POOprod{T_3}{T_4}}}
    : \POOprod{T_3}{T_4}
  }
\\ \\
  \inference{
    v_1 \approx v_2 : T_1
  }{
    \hcvOOinl{v_1}{id(T_1)} \approx \vOOinl{v_2}
    : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
    \hcvOOinl{v_1}{c} \approx v_2
    : \POOsum{T_1}{T_2} &
    c' = seq(c,cast(T_1,l,T_3))
  }{
    \hcvOOinl{v_1}{c'} \approx
    \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
    : \POOsum{T_3}{T_4} 
  }
  \\
  \inference{
  v_1 \approx v_2 : T_2
  }{
  \hcvOOinr{v_1}{id(T_2)} \approx \vOOinr{v_2}
  : \POOsum{T_1}{T_2}
  }
  \quad
  \inference{
  \hcvOOinr{v_1}{c} \approx v_2
  : \POOsum{T_1}{T_2} &
  c' = seq(c,cast(T_2,l,T_4))
  }{
  \hcvOOinr{v_1}{c'} \approx
  \vOOcast{v_2}{\cOOcast{\POOsum{T_1}{T_2}}{l}{\POOsum{T_3}{T_4}}}
  : \POOsum{T_3}{T_4}
  }
  \end{gather*}
  \caption{Bisimulation between \effCEK{C} and \ineffCEKD}
  \label{fig:bisim-SC-D}
\end{figure}
